---
title: "Prestudy Report"
author: "sasha popov"
date: "`r Sys.Date()`"
output:
  rmdformats::html_clean:
    keep_md: true
    self_contained: true
    thumbnails: false
    lightbox: true
    gallery: false
    highlight: tango
    code_folding: hide
    toc: TRUE
    toc_depth: 3
    fig_width: 12
    fig_height: 8

---

## 1. Demographic Data (Exploratry Data Analysis, a.k.a. Overview)

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(gt)
source("analysfunc.R")
df_act = compose_action_df()
resp = elicit_respondents(df_act, no_timestamp = FALSE) |> 
  left_join(institute_lookup, by = c("subj" = "dep")) |> 
  na.omit()
```

Overall participants:

```{r}
print(nrow(resp))
```


Our respondent's data has 6 key observations:  

* `age` - age. 
* `gender` (with ability to enter gender if non-binary). 
* `subj` - department of the student's program/major. 
* `edu_year` - amount of years in the university. 
* `occup` - amount of years in Norway. 
* `family_acad` - highest degree of parents. 
* `orient`- sexual orientation. 

```{r}
resp |>  head() |>  gt()
```

Let's plot demographic data

### Age & Gender

```{r}
resp |> 
  group_by(gender) |> 
  count()
```

We see that gender is evenly distributed over the survey.

```{r}
data_age_gender = resp |> 
    mutate(age = as.numeric(age)) |>
    count(age, gender) |>
    complete(age = 18:40, fill = list(n = 0))

ggplot(data_age_gender) +
  geom_bar(aes(x = age, y = n, fill = gender), stat = "identity") +
  theme_minimal() +
  labs(
    title = "Age/Gender Distribution of Prestudy Respondents",
    subtitle = "absolute values",
    x = "Age",
    y = "Respondents"
  ) +
  scale_fill_manual(values=c("#DD6C40", "#509DA9", "#C4E059"))
```

```{r}
ggplot(data_age_gender |> group_by(age) |> mutate(total_resp = sum(n)) |> filter(total_resp >= 10))+
  geom_bar(aes(x = age, y = n, fill = gender), stat = "identity", position = "fill") +
  theme_minimal() +
  labs(
    title = "Age/Gender Distribution of Prestudy Respondents",
    subtitle = "relative values for years with 10+ obs",
    x = "Age",
    y = "Respondents"
  ) +
  scale_fill_manual(values=c("#DD6C40", "#509DA9", "#C4E059"))
```

Maybe it's make sense to replace gender "other" to non-binary in plots:

```{r}
resp |> 
  filter(gender == "other") |> 
  select(gender_oth) |> 
  head()
```

### Departments & Institutes

We still have majority of students from ex-MATNAT, and 0 respondents from Medicine and Law Faculties:



```{r}
data_inst_dep = resp |> 
  group_by(institute, subj) |> 
  count() |> 
  full_join(institute_lookup, by = c("institute", "subj" = "dep")) |> 
  mutate(n = ifelse(is.na(n), 0, n))

cutiepie = c("Faculty of Fine Art, Music and Design" = "#9A3FE5", "Faculty of Humanities" = "#D5CC4A", "Faculty of Law" = "#B73737", "Faculty of Mathematics and Natural Sciences" = "#2C968C", "Faculty of Medicine" = "#55DB46", "Faculty of Psychology" = "#3E63B3", "Faculty of Social Sciences" = "#C48E43")


institue_color = data_inst_dep |>
  group_by(institute) |> 
  count() |> 
  cbind(color = cutiepie) |>  # 7 institutes 7 colors
  select(-n)

dep_color_palette_vector = generate_department_colors(data_inst_dep, institue_color) 
dep_color_palette = data.frame(Name = names(dep_color_palette_vector), Value = dep_color_palette_vector, stringsAsFactors = FALSE, row.names = NULL)
data_inst_dep = data_inst_dep |> 
  left_join(dep_color_palette, by = c("subj" = "Name")) |> 
  mutate(institute_subj = interaction(institute, subj, sep = ": ")) |> 
  arrange(institute_subj)
```

```{r}
data_inst_dep  |> 
  ggplot() +
  geom_bar(aes(x = institute, y = n), stat = "identity") +
  theme_minimal() + 
  labs(
    title = "Respondents by Institute",
    fill = "Department",
    x = "institute",
    y = "respondents"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```

```{r}
data_inst_dep_nonull = data_inst_dep  |> 
  filter(n>0)
data_inst_dep_nonull |> 
  ggplot() +
  geom_bar(aes(x = institute, y = n, fill = institute_subj), stat = "identity") +
  theme_minimal() + 
  labs(
    title = "Respondents by Department",
    fill = "Department",
    x = "institute",
    y = "respondents"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) + 
  scale_fill_manual(values = data_inst_dep_nonull$Value) + 
  guides(fill = guide_legend(ncol = 1))
```

### Education Year

```{r}
data_edu_years = resp |> 
  mutate(edu_year = as.numeric(edu_year)) |> 
  group_by(gender, edu_year) |> 
  count(edu_year) |> 
  filter(edu_year < 10)

ggplot(data_edu_years) +
  geom_bar(aes(x = edu_year, y = n, fill = gender), stat = "identity") +
  theme_minimal() +
  labs(
    title = "Current Year as a Student",
    subtitle = "Distribution of Prestudy Respondents",
    x = "Years of Education",
    y = "Respondents"
  ) +
  scale_fill_manual(values=c("#DD6C40", "#509DA9", "#C4E059"))
```

### Occupation

This was a proxy on native / international student:

```{r}
cat(sum(resp$age == resp$occup), " ot of ", nrow(resp), " respondents lived in Norway the whole life (", round(sum(resp$age == resp$occup)/nrow(resp)*100,2), "%)", sep = "")
```

Only a small part of students are from abroad: 

```{r}
resp_no_full_norw = resp |> 
  filter(age != occup) |> 
  select(age, occup) |> 
  mutate(
    age = as.numeric(age),
    occup = as.numeric(occup),
    diff = age - occup) |> 
  filter(diff > 0) 

resp_no_full_norw |> 
  summarise(average_age_occup_differnece = mean(diff)) |> 
  gt()
```

```{r}
ggplot(resp_no_full_norw) + 
  geom_bar(aes(x = occup)) + 
  theme_minimal() + 
  labs(
    title = "Years in Norway",
    subtitle = "Distribution for students whos age != occup_years",
    x = "Years in Norway",
    y = "Respondents"
  )
```

### First of the Family

Higher degree of family members can also correlate with the preformance, well-being, etc. 

```{r}
data_family =  resp |> 
  group_by(family_acad, institute) |> 
  count() |> 
  mutate(family_acad = factor(family_acad,
                                 levels = c(
                                   "elementary school",
                                   "secondary school" ,
                                   "Bachelor (or equivalent)",
                                   "Master (or equivalent)",
                                   "PhD")),
         institute = factor(institute, 
                            levels = c("Faculty of Fine Art, Music and Design ", "Faculty of Humanities", "Faculty of Mathematics and Natural Sciences", "Faculty of Psychology", "Faculty of Social Sciences"))) |> 
  left_join(institue_color, by = "institute")

ggplot(data_family) +
  geom_bar(aes(x = family_acad, y = n), stat = "identity") + 
  theme_minimal() + 
  labs(
    title = "Parents' highest degree",
    x = "Highest Degree",
    y = "Respondents"
  ) 
```

```{r}
data_family  |> 
  arrange(institute) |> 
  ggplot() +
    geom_bar(aes(x = family_acad, y = n, fill = institute), stat = "identity") + 
    theme_minimal() + 
    labs(
      title = "Parents' highest degree",
      x = "Highest Degree",
      y = "Respondents", 
      fill = "Faculty"
    ) +
  scale_fill_manual(values = c(cutiepie[-3][-4], "black", "black", "black", "black"))


```

### Sexual Orientation

```{r}
data_so = resp |> 
  filter(orient != "") |> 
  group_by(orient) |> 
  count()

ggplot(data_so) +
  geom_bar(aes(x = orient, y = n), stat = "identity") + 
  theme_minimal() + 
  labs(
    title = "Sexual Orientation",
    x = "Highest Degree",
    y = "Orientation"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```

